---
title: "Domain_analysis"
author: "Shiori"
date: "2026-01-21"
output: html_document
---

```{r}
library(dplyr)
library(tidyr)
library(stringr)
library(tibble)
library(openxlsx)
library(ggplot2)
library(venndir)
library(ggbeeswarm)
library(ggpubr)
library(rstatix)
library(stats)
library(bio3d)
library(httr)
library(jsonlite)
library(R.utils)

source("../Rscript/utils_Domain_analysis.R")

```

# Perform domain enrichment analysis for molecules carrying CFB glycans with altered HexNAc branching
# Use all GNs detected in the synaptosome fraction as the background set
# Use the 33 GNs classified above as showing HexNAc alterations among CFB-glycan-containing molecules as the target set
```{r, fig.width=10, fig.height=5}
dir<-"../demo_data/expected_output/"
protein_db_dir<-"../demo_data/input/"
output_dir_fig<-paste0(dir,"fig/")

sps_gpep <- read.xlsx(paste0(dir,"sps_pattern_result.xlsx"), "pattern_result")

sps_gpep<-rename(sps_gpep, "acc" = "First_acc")

decre_CFBHN_GN<-read.xlsx(paste(dir, "for_strings_sps_HexNAc_decre_GN.xlsx"))%>%pull(decre_CFBHN_GN)

domain_res<-domain_enrichment(decre_CFBHN_GN,sps_gpep)

DoE_plot<-domain_enrich_plot(domain_res[["res"]])

saveRDS(DoE_plot, paste0(output_dir_fig, "domain_enrichment.rds"))
DoE_plot

# view(domain_res[["target_feature"]])
# view(domain_res[["res"]])

```

# Map domain and category information
# Retrieve UniProt domain features and transmembrane regions for altered molecules
# Extract the first transmembrane region
# Filter glycosylation sites in altered molecules:
# retain potential sites from slash-separated site annotations when available;
# otherwise keep the original site annotation
# Extract the first retained site
```{r, fig.width=10, fig.height=5}

sps_ncor_gpep<-read.xlsx(paste0(dir, "CC_sps_noncor.xlsx"), 1)
protein_length<-read.csv(paste0(protein_db_dir, "Swissprot_Mouse_20201124_ver2.csv"))%>%
                select(acc_num,GN, LEN)%>%
                rename("acc" = "acc_num")


decre_glycan<-c("Hex(3)HexNAc(5)Fuc(1)", "Hex(4)HexNAc(5)Fuc(2)")
incre_glycan<-c("Hex(3)HexNAc(4)Fuc(1)", "Hex(4)HexNAc(4)Fuc(2)")

#Cluster 2&3 = increased, 1 = decreased
sps_CFB_incre_df<-filter(sps_ncor_gpep, ratio_cluster %in% c(2,3), glycan_comp %in% incre_glycan, Category == "CF_Bisect")
sps_CFB_decre_df<-filter(sps_ncor_gpep, ratio_cluster %in% 1, glycan_comp %in% decre_glycan, Category == "CF_Bisect")

match_site_list<-list( "Increased_341_442_site" = unique(sps_CFB_incre_df$acc_site_ID), "Decreased_351_452_site" = unique(sps_CFB_decre_df$acc_site_ID))
CFB_site_venn<-venn_analysis(match_site_list)
CFB_match_site<-extract_items(CFB_site_venn@label_df, "Increased_341_442_site&Decreased_351_452_site")%>%unname()



for_strings<-filter(sps_ncor_gpep, acc_site_ID %in% CFB_match_site)%>%
             pull(GN)%>% unique()%>%data.frame()

# write.xlsx(for_strings, paste0(output_dir,"CFB_match_site_GN.xlsx") )

#上chankのCFBの同一サイト上での変化を起こしている部位をマーク
#plotするdataの選別


#domainの抽出
target_GN_df<-filter(sps_ncor_gpep, GN %in% decre_CFBHN_GN)
target_domain<-get_feature(target_GN_df)%>%
               mutate(description = ifelse(description=="", type, description))%>%
               left_join(protein_length, by = "acc")


domain_all<-domain_plot(sps_ncor_gpep, target_domain, "All", CFB_match_site, unique(target_domain$GN))

saveRDS(domain_all, paste0(output_dir_fig, "Glycan_remodeling_domain_map.rds"))
domain_all

```
