---
title: "HexNAc_freq_analysis"
author: "Shiori"
date: "2026-01-21"
output: html_document
---

```{r}
library(dplyr)
library(tidyr)
library(stringr)
library(tibble)
library(openxlsx)
library(ggplot2)
library(venndir)
library(ggbeeswarm)
library(ggpubr)
library(rstatix)
library(stats)
library(bio3d)
library(httr)
library(jsonlite)

source("../Rscript/utils_HexNAc_freq_analysis.R")
```

# Extract GNs with altered HexNAc branching or terminal fucosylation
# Calculate the frequency of HexNAc numbers for each temporal pattern
# Select molecules with a pattern-specific frequency greater than 0.4
# Exclude molecules assigned to two or more patterns because of dispersed distributions
# Identify molecules with a leftward (decreased) or rightward (increased) shift in the HexNAc centroid
```{r, fig.width=10, fig.height=5}

dir<-"../demo_data/expected_output/"
output_dir_fig<-paste0(dir,"fig/")

sps_ncor_gpep<-read.xlsx(paste0(dir, "CC_sps_noncor.xlsx"), 1)
sps_HN_all<-Glycan_class(sps_ncor_gpep, "Branch_HexNAc")
sps_Bisect_HN<-Glycan_class(sps_ncor_gpep, "Bisecting")%>%
               Glycan_class("Branch_HexNAc")
sps_CF_HN<-Glycan_class(sps_ncor_gpep, "CoreFucose")%>%
           Glycan_class("Branch_HexNAc")


all_count_HN_cl<-plot_data_make(sps_HN_all, col = c("acc_site_ID", "pattern"), FALSE, "Branch_HexNAc")%>%
                 group_by(GN) %>%filter(n_distinct(pattern) > 1) %>%ungroup()

CFB_count_HN_cl<-filter(sps_HN_all, Category == "CF_Bisect")%>%
                 plot_data_make( col = c("acc_site_ID", "pattern"), FALSE, "Branch_HexNAc")%>%
                 group_by(GN) %>%filter(n_distinct(pattern) > 1) %>%ungroup()

nCFB_count_HN_cl<-filter(sps_HN_all, Category != "CF_Bisect")%>%
                 plot_data_make( col = c("acc_site_ID", "pattern"), FALSE, "Branch_HexNAc")%>%
                 group_by(GN) %>%filter(n_distinct(pattern) > 1) %>%ungroup()

Bisect_count_HN_cl<-filter(sps_Bisect_HN, all_Bisect !=0)%>%
                    plot_data_make(col = c("acc_site_ID", "pattern"), FALSE, "Branch_HexNAc")%>%
                    group_by(GN) %>%filter(n_distinct(pattern) > 1) %>%ungroup()

non_Bisect_count_HN_cl<-filter(sps_Bisect_HN, all_Bisect ==0)%>%
                        plot_data_make(col = c("acc_site_ID", "pattern"), FALSE, "Branch_HexNAc")%>%
                        group_by(GN) %>%filter(n_distinct(pattern) > 1) %>%ungroup()

CF_count_HN_cl<-filter(sps_CF_HN, all_CF !=0)%>%
                        plot_data_make(col = c("acc_site_ID", "pattern"), FALSE, "Branch_HexNAc")%>%
                        group_by(GN) %>%filter(n_distinct(pattern) > 1) %>%ungroup()

onlyCF_count_HN_cl<-filter(sps_CF_HN, Category == "CF_noBisect")%>%
                        plot_data_make(col = c("acc_site_ID", "pattern"), FALSE, "Branch_HexNAc")%>%
                        group_by(GN) %>%filter(n_distinct(pattern) > 1) %>%ungroup()

onlyB_count_HN_cl<-filter(sps_Bisect_HN, Category == "noCF_Bisect")%>%
                        plot_data_make(col = c("acc_site_ID", "pattern"), FALSE, "Branch_HexNAc")%>%
                        group_by(GN) %>%filter(n_distinct(pattern) > 1) %>%ungroup()

nofe_count_HN_cl<-filter(sps_CF_HN, Category == "noCF_noBisect")%>%
                         plot_data_make(col = c("acc_site_ID", "pattern"), FALSE, "Branch_HexNAc")%>%
                         group_by(GN) %>%filter(n_distinct(pattern) > 1) %>%ungroup()


non_CF_count_HN_cl<-filter(sps_CF_HN, all_CF ==0)%>%
                        plot_data_make(col = c("acc_site_ID", "pattern"), FALSE, "Branch_HexNAc")%>%
                        group_by(GN) %>%filter(n_distinct(pattern) > 1) %>%ungroup()

# Calculate frequencies by HexNAc branch number among altered acc_site_glycan entries
# in GNs with altered HexNAc branching or terminal fucosylation

HN_freq_all<-freq_cal(all_count_HN_cl, "Branch_HexNAc")
B_HN_freq_all<-freq_cal(Bisect_count_HN_cl, "Branch_HexNAc")
nB_HN_freq_all<-freq_cal(non_Bisect_count_HN_cl, "Branch_HexNAc")

CF_HN_freq_all<-freq_cal(CF_count_HN_cl, "Branch_HexNAc")
nCF_HN_freq_all<-freq_cal(non_CF_count_HN_cl, "Branch_HexNAc")

only_CF_HN_freq_all<-freq_cal(onlyCF_count_HN_cl, "Branch_HexNAc")
only_B_HN_freq_all<-freq_cal(onlyB_count_HN_cl, "Branch_HexNAc")

nofe_HN_freq_all<-freq_cal(nofe_count_HN_cl, "Branch_HexNAc")

CFB_HN_freq_all<-freq_cal(CFB_count_HN_cl, "Branch_HexNAc")
nCFB_HN_freq_all<-freq_cal(nCFB_count_HN_cl, "Branch_HexNAc")

# Extract molecules with a frequency of 0.4 or higher to exclude molecules in which increasing and decreasing patterns are dispersed and therefore show little apparent directional change 
#Note: whether 0.4 is an appropriate threshold should be further evaluated
# Process molecules with three or more rows within the same GN: for entries with the same HexNAc (or terminal Fuc) number, compare their frequencies and retain the entry with the higher frequency
# Because increasing and decreasing patterns coexist at the same HexNAc number, evaluate which direction of change is more significant
# When increasing and decreasing patterns are distributed across different HexNAc
# or terminal Fuc numbers, calculate the mean HexNAc number for increasing entries
# or decreasing entries, and subtract the corresponding reference value

extract_HN_change_GN<-glycan_dif_analysis(HN_freq_all, "Branch_HexNAc")
extract_BHN_change_GN<-glycan_dif_analysis(B_HN_freq_all, "Branch_HexNAc")
extract_nBHN_change_GN<-glycan_dif_analysis(nB_HN_freq_all, "Branch_HexNAc")
extract_CFHN_change_GN<-glycan_dif_analysis(CF_HN_freq_all, "Branch_HexNAc")
extract_nCFHN_change_GN<-glycan_dif_analysis(nCF_HN_freq_all, "Branch_HexNAc")
extract_CFBHN_change_GN<-glycan_dif_analysis(CFB_HN_freq_all, "Branch_HexNAc")
extract_nCFBHN_change_GN<-glycan_dif_analysis(nCFB_HN_freq_all, "Branch_HexNAc")
extract_onCFHN_change_GN<-glycan_dif_analysis(only_CF_HN_freq_all, "Branch_HexNAc")
extract_onBHN_change_GN<-glycan_dif_analysis(only_B_HN_freq_all, "Branch_HexNAc")
extract_nofe_change_GN<-glycan_dif_analysis(nofe_HN_freq_all, "Branch_HexNAc")

extract_CFBHN_change_GN<-glycan_dif_analysis(CFB_HN_freq_all, "Branch_HexNAc")

decre_HN_GN<-extract_HN_change_GN%>%filter(dif_num<0)%>%pull(GN)%>%unique()
incre_HN_GN<-extract_HN_change_GN%>%filter(dif_num>0)%>%pull(GN)%>%unique()

decre_BHN_GN<-extract_BHN_change_GN%>%filter(dif_num<0)%>%pull(GN)%>%unique()
incre_BHN_GN<-extract_BHN_change_GN%>%filter(dif_num>0)%>%pull(GN)%>%unique()

decre_nBHN_GN<-extract_nBHN_change_GN%>%filter(dif_num<0)%>%pull(GN)%>%unique()
incre_nBHN_GN<-extract_nBHN_change_GN%>%filter(dif_num>0)%>%pull(GN)%>%unique()

decre_CFHN_GN<-extract_CFHN_change_GN%>%filter(dif_num<0)%>%pull(GN)%>%unique()
incre_CFHN_GN<-extract_CFHN_change_GN%>%filter(dif_num>0)%>%pull(GN)%>%unique()

decre_nCFHN_GN<-extract_nCFHN_change_GN%>%filter(dif_num<0)%>%pull(GN)%>%unique()
incre_nCFHN_GN<-extract_nCFHN_change_GN%>%filter(dif_num>0)%>%pull(GN)%>%unique()

decre_CFBHN_GN<-extract_CFBHN_change_GN%>%filter(dif_num<0)%>%pull(GN)%>%unique()
incre_CFBHN_GN<-extract_CFBHN_change_GN%>%filter(dif_num>0)%>%pull(GN)%>%unique()

decre_nCFBHN_GN<-extract_nCFBHN_change_GN%>%filter(dif_num<0)%>%pull(GN)%>%unique()
incre_nCFBHN_GN<-extract_nCFBHN_change_GN%>%filter(dif_num>0)%>%pull(GN)%>%unique()

decre_onCFHN_GN<-extract_onCFHN_change_GN%>%filter(dif_num<0)%>%pull(GN)%>%unique()
decre_onBHN_GN<-extract_onBHN_change_GN%>%filter(dif_num<0)%>%pull(GN)%>%unique()

decre_nofe_GN<-extract_nofe_change_GN%>%filter(dif_num<0)%>%pull(GN)%>%unique()


decre_CFBHN_GN<-extract_CFBHN_change_GN%>%filter(dif_num<0)%>%pull(GN)%>%unique()
incre_CFBHN_GN<-extract_CFBHN_change_GN%>%filter(dif_num>0)%>%pull(GN)%>%unique()

           
decre_CFHN_pair<-filter(CF_count_HN_cl, GN %in% decre_CFHN_GN)%>%
               pairwise_res("Branch_HexNAc", TRUE)

decre_BHN_pair<-filter(Bisect_count_HN_cl, GN %in% decre_BHN_GN)%>%
               pairwise_res("Branch_HexNAc", TRUE)

decre_CFBHN_pair<-filter(CFB_count_HN_cl, GN %in% decre_CFBHN_GN)%>%
               pairwise_res("Branch_HexNAc", TRUE)

decre_onCFHN_pair<-filter(onlyCF_count_HN_cl, GN %in% decre_onCFHN_GN)%>%
               pairwise_res("Branch_HexNAc", TRUE)

decre_nofe_pair<-filter(nofe_count_HN_cl, GN %in% decre_nofe_GN)%>%
               pairwise_res("Branch_HexNAc", TRUE)

decre_onBHN_pair<-filter(onlyB_count_HN_cl, GN %in% decre_onBHN_GN)%>%
               pairwise_res("Branch_HexNAc", TRUE)

pair_wise_output_list<-list("CF(+CFB)_glycan_pairwise"=decre_CFHN_pair,
                            "B(+CFB)_glycan_pairwise"=decre_BHN_pair,
                            "CFB_glycan_pairwise"=decre_CFBHN_pair,
                            "only_CF_glycan_pairwise"=decre_onCFHN_pair,
                            "no_feature_glycan_pairwise"=decre_nofe_pair)

output_dir<-dir
write.xlsx(pair_wise_output_list, paste0(output_dir, "HN_decrease_pair_wise_res.xlsx"))


freq_HN_decre<-filter(HN_freq_all, GN %in% decre_HN_GN)%>%
               count_heatmap("Branch_HexNAc", "GN", FALSE)

freq_CFHN_decre<-filter(CF_HN_freq_all, GN %in% decre_CFHN_GN)%>%
               count_heatmap("Branch_HexNAc", "GN", FALSE)

freq_BHN_decre<-filter(B_HN_freq_all, GN %in% decre_BHN_GN)%>%
               count_heatmap("Branch_HexNAc", "GN", FALSE)

freq_onCF_decre<-filter(only_CF_HN_freq_all, GN %in% decre_onCFHN_GN)%>%
               count_heatmap("Branch_HexNAc", "GN", FALSE)

freq_onBHN_decre<-filter(only_B_HN_freq_all, GN %in% decre_onBHN_GN)%>%
               count_heatmap("Branch_HexNAc", "GN", FALSE)

freq_nofeHN_decre<-filter(nofe_HN_freq_all, GN %in% decre_nofe_GN)%>%
               count_heatmap("Branch_HexNAc", "GN", FALSE)

freq_nBHN_decre<-filter(nB_HN_freq_all, GN %in% decre_BHN_GN)%>%
               count_heatmap("Branch_HexNAc", "GN", FALSE)

freq_nCFHN_decre<-filter(nCF_HN_freq_all, GN %in% decre_nCFHN_GN)%>%
               count_heatmap("Branch_HexNAc", "GN", FALSE)

freq_CFBHN_decre<-filter(CFB_HN_freq_all, GN %in% decre_CFBHN_GN)%>%
               count_heatmap("Branch_HexNAc", "GN", FALSE)

match_protein<-list("CF" = decre_CFHN_GN, "Bisect" = decre_BHN_GN, "CFB" = decre_CFBHN_GN)
match_protein_venn<-venn_analysis(match_protein)

saveRDS(match_protein_venn, paste0(output_dir_fig, "HexNAc_decre_venn.rds"))

for_strings_data<-data.frame(decre_CFBHN_GN)
write.xlsx(for_strings_data, paste(output_dir, "for_strings_sps_HexNAc_decre_GN.xlsx"))

saveRDS(freq_HN_decre, paste0(output_dir_fig, "freq_HN_decre.rds"))
saveRDS(freq_CFHN_decre, paste0(output_dir_fig, "freq_CFHN_decre.rds"))
saveRDS(freq_BHN_decre, paste0(output_dir_fig, "freq_BHN_decre.rds"))
saveRDS(freq_onCF_decre, paste0(output_dir_fig, "freq_onCF_decre.rds"))
saveRDS(freq_onBHN_decre, paste0(output_dir_fig, "freq_onBHN_decre.rds"))
saveRDS(freq_nofeHN_decre, paste0(output_dir_fig, "freq_nofeHN_decre.rds"))
saveRDS(freq_nBHN_decre, paste0(output_dir_fig, "freq_nBHN_decre.rds"))
saveRDS(freq_nCFHN_decre, paste0(output_dir_fig, "freq_nCFHN_decre.rds"))
saveRDS(freq_CFBHN_decre, paste0(output_dir_fig, "freq_CFBHN_decre.rds"))

```